Benchmarking neuromorphic systems with Nengo

### 2015-09-13 paper/rebuttals.txt

Fair point. For logistical reasons, it is often nice to be able
to set a seed and track it; for example, in collecting data for
this paper, each model's results are stored in a file whose name
includes the seed. This makes it easy to ensure that the same
model is not run twice, biasing the sample.
However, in general, it is true that setting seeds is not important
for benchmarking. For this reason, we have removed the mentions
of setting seeds, including references to the `rng` and `seed`
fixtures. We have kept one reference in the Background section discussing
the frontend interface, but have  reworded that it slightly;
it now reads:

  The only parameter on the network is an optional integer seed;
  setting this should make all of the random factors in the model
  deterministic, which is important for testing and debugging models,
  but is not used for benchmarking.

> - The authors state that Nengo is well-suited for generating large
>   networks. However, the tests they present are only limited in
>   size. A comment on larger benchmark models in the test suite
>   and/or the possibility to scale up the present benchmarks would be
>   desirable.

We have added an explicit mention of the size of Spaun on lines 401-402
(since Spaun is built with Nengo and shows that it is possible
to generate large networks). We have also added a short paragraph
to the Test models subsection discussing larger models;
see lines 255-259 which read:

  We have used relatively small models run for short times in order to
  run many iterations on all backends.  However, all of these models
  could be made significantly larger by increasing the number of
  neurons used (i.e., adjusting the \texttt{n\_neurons} parameter on
  ensembles), and increasing the dimensionality of the signals
  represented in the model.  Additionally, existing tests of the
  semantic pointer architecture within Nengo use significantly more
  neurons than the models presented here, and could be adapted into
  benchmarks.

> Minor:
>
> - I found the heading of section 2.2 ("Fuzzy testing") puzzling - is
>   this a fixed term I am missing? In any case, it would be helpful
>   to explain the term.

The term "Fuzzy testing" borrows from the concept of Fuzzy logic
from computer science. However, upon searching for other uses
of the term, we noticed that the term "Fuzz testing" is commonly
used to refer to automated testing with random data.
In order to avoid confusion with "Fuzz testing",
and to avoid inventing new parlance for litte gain,
we have changed the heading of section 2.2 to "Functional testing",
and have removed other references to "fuzzy testing".

> - Table 2 is referenced before Table 1 in the text - their order
>   should be changed

Thank you, this has been corrected.

@@

5. Other comments on the results.
---------------------------------

> I found the description of the Figures on accuracy too short. Even
> if the accuracy results themselves are not the main focus of the
> article, some more explanations for the results should be given. In
> particular, in Figs. 1 and 2, why does the Nengo reference backend
> show a significantly higher variability than the Brainstorm and
> distilled backends? In Figs. 3 and 4, where do the deviations in the
> SpiNNaker backend come from (my first guess would be the employed
> 1ms time step, which is larger than the 0.1ms typically used in
> software simulators)?

We have added additional details in the section 4.2 on accuracy.
These additional details are listed below.

We are still in the progress of investigating the deviations
in the SpiNNaker backend; however, we are collaborating with
the SpiNNaker group to improve accuracy on these examples.
The 1ms time step is not the issue here, however,
as all of the backends tested use a 1ms time step.

Figure 1:

  The Brainstorm and distilled backends have the least variability and
  the SpiNNaker backend has the most variability, though the median
  RSME is the same across all backends.  One driver of these
  differences is in how each backend implement synaptic filtering.
  The reference and OpenCL backends, for example, implement arbitrary
  linear time-invariant filters, of which the lowpass filter used in
  this model is a specific example.  The Brianstorm and distilled
  backends only support lowpass filters, and therefore implement it
  more accurately.  Another reason for these differences is in how
  each backend handles ensembles operating near the edge of their
  representational range (i.e., the radius of the ensemble).  The
  outlier with highest RMSE represents a model instance in which the
  static vector target was at the extreme of the representational
  range; the SpiNNaker backend's relatively high RMSE on this example
  indicates that it may not perform as well as other backends in this
  situation.  However, accuracy could be improved by increasing the
  representational range, though doing so may necessitate an increase
  in the number of evaluation points generated when solving for
  decoding weights.

Figure 2:

  Again, the more generic synaptic filtering model used in the
  reference and OpenCL backends is the likely cause of increased
  variability on this model.  It is also likely that performance at
  the extremes of the representational range is responsible for the
  SpiNNaker backend's reduced accuracy, though we have not
  investigated this in detail.

Figure 3:

  In this example, differences in synaptic filtering are more
  pronounced than in previous examples, as the recurrent connection is
  responsible for the dynamics of the oscillation.  This result
  indicates that the synaptic filtering model of SpiNNaker may also
  warrant closer examination for accuracy.

Figure 4:

  The large difference between the two versions of this model for the
  SpiNNaker backend indicates that node-to-ensemble and
  ensemble-to-node connections introduce additional delays that are
  not present on other backends.  This result is to be expected, as
  nodes can execute arbitrary code and are therefore difficult to
  simulate in real time with special purpose hardware.

> Also, the whiskers in Figs. 1-4 seem to be only multiples of the box
> sizes, so plotting them does not add any information - it may be
> even misleading when not carefully reading the caption.

Thank you for bringing up this point!
Our description of the box plot was not quite correct
in the initial submission. We have rewritten this description
as follows:

  The box shows the median and interquartile range, while the whiskers
  extend to the farthest data point within 1.5 times the interquartile
  range; outliers are shown above and below the whiskers.

As noted in these captions, we have also now included the outliers.
We were initially concerned that outliers would make the rest of the
boxplot too small relative to the expanded axes caused by outliers.
However, we now feel that including all of the data makes these
boxplots more useful.

> Some of the results in Fig. 7 are puzzling: For BG sequence*, the
> run time for nengo and nengo_ocl is even higher without probes
> (compared to Fig. 6), being well outside the 95% confidence
> intervals. This should be explained.

We did not notice the slowdown upon initial submission,
so seeing this was puzzling to us as well.
To investigate whether this effect was real,
we first replicated it on two additional machines;
the version without probes was significantly slower
than the version with probes on all machines,
as measured by non-overlapping
bootstrapped 95% confidence intervals.
Then, we investigated the internal structures of the two
versions of networks, in terms of the step functions
produced by them.
The version with probes had 600 step functions;
595 of these implemented the same equations
as the networks without probes, while the
remaining five functions implemented the probing.
The simulator loop also indicated that
the version without probes should have executed
strictly fewer operations than the version with probes.
We conclude that the reason for the speedup
must have been somehow related to
a lower level of abstraction than we investigated
(e.g., the data was better organized in RAM
in the version with probes).

Very few networks in Nengo have zero probes,
so there were some obvious opportunities to optimize
the reference backend for the special case
in which no probes exist. After implementing
two such optimizations, we were able to consistently
achieve identical performance with probes and without.
However, it was not our intention to perform
a detailed comparison of the affect of probes
on runtimes, and we feel that including Figure 7
(which only had a few sentences of text associated with it)
was more distracting than illuminating.
For these reasons, we have removed Figure 7 in this revision.

@@

6. Other comments on the discussion or the conclusions.
-------------------------------------------------------

> As already mentioned for the Introduction, it would make the
> Discussion section much more balanced if the limitations of the
> approach would be discussed as well, together with some outlook to
> potential improvements or complementary approaches.

As previously mentioned, we added in a paragraph to the discussion
on the primary weakness of the approach.
This paragraph also includes a potential improvement:
collecting power consumption information (lines 417-423).

> Minor:
>
> - typo in line 373: "We believe that THE these benchmarks..."

Corrected, thank you!

@@

7. Please add here any further comments on this manuscript.
-----------------------------------------------------------

> Overall, the article presents an attractive idea for benchmarking
> neuromorphic systems. In my opinion its main weakness is the lack of
> embedding in the current literature. Apart from that, it is a nice
> contribution to the field, and the open source code is a good
> opportunity for other neuromorphic system operators for benchmarking
> their systems.
>
> Minor:
>
> - missing citation details for Stewart et al. 2010

Corrected, thank you! In correcting this, we noticed that
Mundy et al., 2015 was also missing details,
which have been filled in.
